---
title: "Import_tidy"
author: 'Moran Wang & Jingru Ma '
date: "January 31, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
library(zoo)
library(ggplot2)
library(readr)
source("import_tidy_function.R")
source("BulletGraph.R")
```


## Sevice Learning
```{r}
#read data 
serviceLearning <- redata ("Service Learning.csv")

##subset the FY2017-18 to FY 2015, change header names
SL_data <- sub_range(serviceLearning,4,45,1,9)
colnames(SL_data) <-c("Month","Zone1","Zone2","Zone3","TNStudents","Tours","Classes","SLR","Reunve")
# delete the rows that are not useful 
d_v <- c ('FY 16-17', '2016', 'FY 15-16','2015')
SL_data <- filter (SL_data, !SL_data$Month %in% d_v)
#str(SL_data)

#test function month_format , good 
testnew <- SL_data$Month
mth <- month_format(testnew)

# test function year_format, good
yrnew <- year_format(SL_data$Month)

#try to add the mth , int 2,3,4 with yrnew, factor 2018,2017...
#hh <- as.yearmon(paste(yrnew, mth), "%Y%m", sep="-")
#dates1 <- as.Date(hh)
#SL_data$graphYearCol <- clean_date(SL_data$Month)

#change month from Feburary 18 to Feb 18
SL_data$Month <- as.yearmon(paste(year_format(SL_data$Month), month_format(SL_data$Month)), "%Y%m", sep="-")

#add a new column in SLrevenue
SL_data$SLRIn <- make_int(SL_data$SLR)

#Zone23= zone2 + zone3 
temp <- make_int(SL_data[1:14,3]) + make_int(SL_data[1:14,4])
temp2 <- make_int(SL_data[15:38,4])
SL_data$zone23 <- c(temp,temp2)

##draw graph of year and total service learning reveune
# the graph has already removed the missing value

#if the revenue is 0, i decided to calcualte an average value to replace it
m <- mean(SL_data$SLRIn,na.rm=TRUE)
SL_data$SLRIn[7] <- m

#bullet graph, Service Learning Revenue
#box plot, var 

#split Feb 2018 to 2018
SL_data <- na.omit(SL_data)
SL_data <- mutate(SL_data,year=sapply(str_split(SL_data$Month," "),function(x){x[2]}))
b1 <- boxplot <- ggplot (SL_data, aes(x = SL_data$year, y = SL_data$SLRIn)) + 
  geom_boxplot()+ xlab("Years") + 
  ylab("Service Learning Revenue") + ggtitle("Revenue in Years")

graphx <- max(SL_data$SLRIn, na.rm= TRUE)
bulletgraph(x=graphx,ref=4500,limits=c(0,3500,5000,6000),name= "",subname="(U.S. $ in \n thousands)",width=0.2)
```
The KPI of servise learning, we shows the revenue of service learning graphically. The reference point is 4500 and the three levels of background are 3500,5000 and 6000. The blue bar reach to the green level which is the excellent level.
```{r}

boxplot #looks like revenue have outlier 

# the max point 
boxout=boxplot(SL_data$SLRIn)$out 
outname=as.character(SL_data$SLRIn) 
outname[(SL_data$SLRIn %in% boxout)==FALSE]="\n" 

outname

b1 <- b1+geom_text(label=outname)  # the first is 3346, 
b1 

str(SL_data)

#summary of the boxplot
#create a new data frame, like slr 2016, 2017, 2018
slr16 <- subset(SL_data, SL_data$year == 2016)
slr17 <- subset(SL_data, SL_data$year == 2017)
slr15 <- subset(SL_data, SL_data$year == 2015)

summary(slr15$SLRIn)
summary(slr16$SLRIn)
summary(slr17$SLRIn)

# graph for revenue by yearly to see it changes, waterfall graph
#covert char to number
SL_data$year <- as.numeric(SL_data$year)

f1 <- aggregate (SLRIn~year, SL_data, sum)
f1$id <- seq_along(f1$year)
diff2 <- f1[2, 2] - f1[1,2 ]
diff3 <- f1[3, 2] - f1[2,2 ]

f1$diff <- c(13759.3, diff2, diff3)

f1$end <- c (13759.30, 20769.20,13479.04)
f1$start <- c(0, head(f1$end, -1))
f1$end <- c(head(f1$end, -1), 13479.04)

f1$year <- as.factor(f1$year)
f1$direction <- ifelse(f1$diff > 0, "Positive", "Negative")

f1 

p1 <- ggplot(f1, aes(f1$year, fill = f1$direction)) + geom_rect(aes(f1$year,xmin = f1$id - 0.45, xmax = f1$id + 0.45, ymin = f1$end, ymax = f1$start)) + xlab ("Year") + labs(title = "Revenue Change by Year")

p1


# monthly revenue change
g <- ggplot (SL_data, aes (SL_data$Month,SL_data$SLRIn)) + geom_line()  + xlab(" ") + ylab("Service Learning Revenue")
g





```
This graph is showing the service learning revenue in the year of 2015,2016 and 2017 as a boxplot. The maximum, third quartile,median, first quartile and minimum of data are showing on the plot. 
```{r}

#analysis: fit linear regression model, the zone1+ zone2+zone3(TNstudents)
lr_data <- SL_data[,c(5,10)]
lr_data$TNStudents <- make_int(lr_data$TNStudents)
new.data <- lr_data[rowSums(is.na(lr_data[ ,c("TNStudents", "SLRIn")])) < 2, ]

# linear regression 
p1 <- ggplot (data = new.data, aes(x= TNStudents, y = SLRIn)) + geom_point() + xlab("Student") + 
  ylab("Revenue") + ggtitle("Total#Students VS Revenue")
p1

```
This graph is showing the relationship of students and revenue. The data of students is independent variables, the data of revenue is dependent variables. 
```{r}

lm1 <- lm(new.data$SLRIn ~ new.data$TNStudents)
m <- 2.686 
b <- 991.115
g <- p1 + geom_abline(slope = m, intercept = b)
g

summary(lm1)

# residual
lm1$residuals

# test if residual normal distrbution
shapiro.test(lm1$residuals)  

par(mfrow=c(2,2))
plot(lm1) #the result is not trusted 


```
Adding a line into graph p1. This graph is showing as a linear regression. The points are not extremely close to the linear. Therefore, the students are revenue might not related to each other.

```{r}
# R- square, residuals
summary(lm1)
# I think linear line is not a good way to fit this relation
```



##Food Box
```{r}
FoodBox <- redata ("FoodBox.csv")
FB_data <- sub_range(FoodBox,2,164,1,8)
colnames(FB_data) <-c("Date","TotalBoxes","BasicBoxes","FamilyBoxes","Zone1","Zone2","Zone3","Revenue")

Month17 <- c('January','February','March','April','May','June', 'July', 'August', 'September','October' , 'November', 'December' )
FB_Month17 <- filter (FB_data, FB_data$Date %in% Month17)

d_v <- c ('FY 16-17', '2016', 'FY 15-16','2015', 'FY 17-18',  'YTD' , 'January','February','March','April','May','June', 'July', 'August', 'September','October' , 'November', 'December')
FB_data <- filter (FB_data, !FB_data$Date %in% d_v)
#str(FB_data)

#convert totalboxes, basic boxes, family boxes, zone1 ,zone2, zone3 to numeric number 
FB_data[2:7] <- lapply(FB_data[2:7], function(x) as.numeric(as.character(x)))

#date is a factor, convert it to a date
FB_data$Date <- as.Date(FB_data$Date, format = "%m/%d/%Y")

FB_data <- na.omit(FB_data)

str(FB_data)

#aggregate the total number of box in different month by year
# how to convert a date to a short date: FB_data$shortdate <- strftime(FB_data$Date, format="%Y/%m")
sb_FB_data <- subset(FB_data[2:7])
sum_year_month <- aggregate(sb_FB_data,by=list((substr(FB_data$Date,1,7))),sum) 
str(sum_year_month) # the group.1 is a charactor , i need to change it to the date to plot 
 
#aggregate_month(FB_data,2:7,FB_data$Date) #this function return the argument must have sum length error

#convert charactor as date
sum_year_month$Group.1 <- as.yearmon(sum_year_month$Group.1)
str(sum_year_month)

#basically, except the 2017-11, the other month's total boxes = bassic boxes + family boxes
variable <- c(sum_year_month$Zone1, sum_year_month$Zone2 , sum_year_month$Zone3)
g <- ggplot (sum_year_month, aes (sum_year_month$Group.1,sum_year_month$Zone1)) + geom_line(aes(color = "Zone1"))  + xlab(" ") + ylab("FoodBoxes")
g <- g + geom_line(aes(y = sum_year_month$Zone2, color = "Zone2")) 
g <- g + geom_line(aes(y = sum_year_month$Zone3, color = "Zone3"))
g <- g + scale_color_manual(values = c("red", "blue", "green")) +
  labs(color="legend") 
g
# for most of the sum zone1, zone2 and zone3, their value equal to total boxes. But for some of them, they are not equal. 
```
This graph is showing the line graphs of Zone1, Zone2 and Zone3 in the Food Boxes dataset. The data of Zone3 are large than the other two.
```{r}
#need to deal with the table contained the revenue through 2015-2017

# try to create a bar graph same with excel, year, zone boxes, zones (like zone1, zone2, zone3)
new_graph <- sum_year_month %>% gather ('Zone1','Zone2','Zone3',key = "Date", value = 'TotalBoxes')
ngraph <- new_graph[c(1,5:6)]

ggplot(data=ngraph, aes(x=Group.1, y=TotalBoxes, fill=ngraph$Date)) +
geom_bar(stat="identity", position=position_dodge()) + 
labs(x="Month", title = "Food Boxes by Zone", fill = "Zone")


#KPI
bulletgraph(x=max(FB_data$TotalBoxes),ref=80,limits=c(0,50,100,150),name= "",subname="(TotalBoxes)",width=0.2)

# check is there any linear relationship between basic box and zone3
m <- ggplot (data = FB_data, aes(x= FB_data$BasicBoxes, y = FB_data$Zone3)) + geom_point() + xlab("basic box") + 
  ylab("zone3") + ggtitle("#of basic food boxes VS #of total boxes in Zone3")+ geom_abline(slope = 0.89559, intercept = 2.51331)

lmn <- lm(FB_data$Zone3 ~ FB_data$BasicBoxes)
summary(lmn)
par(mfrow=c(2,2))
plot(lmn)

shapiro.test(lmn$residuals)


m <- ggplot (data = FB_data, aes(x= FB_data$BasicBoxes, y = FB_data$Zone3)) + geom_point() + xlab("basic box") + 
  ylab("zone3") + ggtitle("Total#Students VS Revenue")
m

m <- ggplot (data = FB_data, aes(x= FB_data$FamilyBoxes, y = FB_data$Zone3)) + geom_point() + xlab("basic box") + 
  ylab("zone3") + ggtitle("Total#Students VS Revenue")
m


```
Showing the KPI of total food boxes. The reference point is 80. The three levels of background are 50,100 and 150. The blue bar reach to the green level which is the excellent level.

## Market
```{r}
Market <- redata ("Market.csv")
Market_data <- sub_range(Market,11,166,1,9)
colnames(Market_data) <-c("Date","Zone1","Zone1Sale","Outside","Zone2","Zone2Sale","Zone3","Zone3Sale", "TotalSales")

d_v <- c ('FY 16-17', '2016', 'FY 15-16','2015', 'FY 17-18',  'YTD' , 'January','February','March','April','May','June', 'July', 'August', 'September','October' , 'November', 'December', 'June total')
Market_data <- filter (Market_data, !Market_data$Date %in% d_v)

#convert factor to number
Market_data[2:9] <- lapply(Market_data[2:9], function(x) as.numeric(as.character(x)))

#convert the factor date to date 
Market_data$Date <- as.Date(Market_data$Date, format = "%m/%d/%Y")

str(Market_data)

#add total by month by year
aggregate_month <- function (data, x, y){
  #x is the range of the data you want to add
  #y is the column of the year month
  temp_data <- subset (data[x])
  sum1 <- aggregate(temp_data,by = list(substr(data$y,1,7)),sum)
  return (sum1)
}

#remove all columns with na 
nmd <- Market_data[ , colSums(is.na(Market_data)) == 0]


g <- ggplot (nmd, aes (nmd$Date,nmd$TotalSales)) + geom_line()  + xlab(" ") + ylab("Total sale")
g
```
This graph is showing the total sales in different date from data. As the graph shows, the total sales are growing basically.
```{r}

#need to read the table includes revenue and total boxes by month
Market_Revenue <- sub_range(Market,4,40,11,13)
#delete blank value
Market_Revenue <- Market_Revenue[-which(Market_Revenue$MONTHLY == ""),]
mr <- Market_Revenue [-20, ]

mr$X.9 <- make_int(mr$X.9)
mr$X.10 <- make_int(mr$X.10)

mr$MONTHLY <- as.Date(mr$MONTHLY, format = "%m/%d/%Y")

g3 <- ggplot (mr, aes (mr$MONTHLY,mr$X.10)) + geom_line(aes(color = "Revenue"))  + xlab(" ") + ylab("Market Revenue")
g3 <- g3 + geom_line(aes(y = mr$X.9, color = "TotalSale"))
g3
```
This graph is comparing the Revenue and total sales of market in different data. The total sale is clearly lower than revenue everyday.

```{r}


#linear regression , total sale and revenue
#density plot, check distribution
p1 <- ggplot (data = mr, aes(x= mr$X.9, y = mr$X.10)) + geom_point() + xlab("TotalSale") + 
  ylab("Revenue") + ggtitle("TotalSale VS Revenue")
p1

lm.sr <- lm(mr$X.10 ~ mr$X.9)
m <- 8.431 
b <- 419.221
g <- p1 + geom_abline(slope = m, intercept = b)
g

# residual
lm.sr$residuals

# test if residual normal distrbution
shapiro.test(lm.sr$residuals) 

data.resid <- bind_cols(mr,fitted = lm.sr$fitted.values)
g<-g + geom_linerange(data=data.resid,aes(x=data.resid$X.9,
                                          ymin=data.resid$X.10,ymax=data.resid$fitted))
g

summary(lm.sr)
par(mfrow=c(2,2))
plot(lm.sr)

```
Adding a linear regression to show the relationship between revenue and total sale. The points are close to the line. Thus, Total sale and revenue are related to each other.
```{r}

# R- square, residuals
summary(lm.sr)
qplot(mr$X.9, lm.sr$residuals) + xlab("total sales")

qqnorm(lm.sr$residuals)
qqline(lm.sr$residuals)

# KPI
bulletgraph(x=max(mr$X.10, na.rm=TRUE),ref=4500,limits=c(0,3500,5500,7000),name= "",subname="(U.S. $ in \n thousands)",width=0.2)
```
Showing the KPI of market revenue. The reference point of this graph is 4500, the three levels of background are 3500,5500 and 7000. The blue bar reach to the green level which is the excellent level. 
